\subsubsection{La-ER (Learning-rate-adaptive Experience Replay)}

\gls{laer} is an ablation variant that decouples the two adaptations introduced by \gls{lamaml}: it uses the same per-parameter learning rate meta-learning to decide \emph{how large each weight step should be}, but replaces the meta-learned weight update with a plain \gls{er} step on the replay buffer. This isolates the contribution of adaptive step-size learning from the full MAML-style weight optimisation.

\paragraph{Memory management}

\gls{laer} maintains a fixed-capacity replay buffer $\mathcal{M}$ of size $K$ updated by reservoir sampling, identical to \gls{er-res} (cf.\ \eqref{eqn:reservoir}). A mixed replay batch $B^\text{rep} = (x^\text{rep}, y^\text{rep}, t^\text{rep})$ is drawn once per observe call and is shared across both the meta-objective and the subsequent weight update.

\paragraph{Per-parameter adaptive learning rates}

As in \gls{lamaml}, the network backbone is augmented with a \texttt{ParameterList} of per-parameter scalars $\hat\alpha_j$ (one per backbone parameter) initialised to $\alpha_\text{init}$ and constrained to be non-negative via a ReLU during application. The rates are updated by a separate SGD optimiser with step size $\eta_\alpha$.

\paragraph{Phase 1: inner update and learning-rate meta-training}

The incoming mini-batch of size $B$ is divided into $K_\text{meta}$ sub-batches. For sub-batch $i$, a fast-weight copy is advanced from the previous sub-batch's fast weights using the same inner step as \gls{lamaml} \eqref{eqn:lamaml_inner}:
%
\begin{equation}\label{eqn:laer_inner}
    \theta'_j
    =
    \theta_j
    -
    \mathrm{ReLU}(\hat\alpha_j) \cdot
    \frac{\partial \mathcal{L}_\text{CE}(B_i;\,\theta)}
         {\partial \theta_j}.
\end{equation}
%
The fast weights are then used to evaluate the current-task loss on the replay batch:
%
\begin{equation}\label{eqn:laer_meta_loss}
    \mathcal{L}_\text{meta}
    =
    \mathcal{L}_\text{CE}\!\left(B^\text{rep};\, \theta'_{K_\text{meta}}\right).
\end{equation}
%
Gradients flow back through the inner step \eqref{eqn:laer_inner} to update only the learning-rate parameters:
%
\begin{equation}\label{eqn:laer_lr_update}
    \hat\alpha_j
    \leftarrow
    \hat\alpha_j
    -
    \eta_\alpha \cdot
    \frac{\partial \mathcal{L}_\text{meta}}{\partial \hat\alpha_j}.
\end{equation}
%
Crucially, the backbone weights $\theta$ are \emph{not} updated in this phase. The meta-gradient only shapes the per-parameter step sizes.

\paragraph{Phase 2: weight update via plain ER}

After the learning-rate update, the backbone weights are updated using a standard replay step. The same replay batch $B^\text{rep}$ is fed through the network at the \emph{current} (non-fast) weights, and the replay cross-entropy loss is computed:
%
\begin{equation}\label{eqn:laer_er_loss}
    \mathcal{L}_\text{ER}
    =
    \mathcal{L}_\text{CE}\!\left(B^\text{rep};\, \theta\right).
\end{equation}
%
The gradient of $\mathcal{L}_\text{ER}$ is then applied using the freshly-updated, meta-informed learning rates as per-parameter step sizes:
%
\begin{equation}\label{eqn:laer_weight_update}
    \theta_j
    \leftarrow
    \theta_j
    -
    \mathrm{ReLU}(\hat\alpha_j) \cdot
    \frac{\partial \mathcal{L}_\text{ER}}{\partial \theta_j}.
\end{equation}

\paragraph{Relationship to La-MAML}

In \gls{lamaml}, the weight update \eqref{eqn:lamaml_weight_update} is driven by the gradient of the \emph{meta-loss} (evaluated with fast weights), so the backbone update direction is shaped by multi-task replay generalisation. In \gls{laer}, the weight update direction comes from plain single-forward ER loss \eqref{eqn:laer_er_loss}, while only the \emph{step sizes} are shaped by the meta-objective. This separation tests whether the benefit of \gls{lamaml} lies in its adaptive per-parameter rates alone, or whether the meta-gradient weight direction is also essential.

\paragraph{Inference}

\gls{laer} uses a standard linear classifier at inference time. In \gls{til}, task identity selects the relevant logit slice. In \gls{cil}, the full logit vector across all seen classes is used.
